TacticToe: Learning to Prove with Tactics

### 2018-08-17 tactictoe_jv.tex
theorem list

\subsection{Abstraction of Tactic Arguments}
The first step is to abstract arguments of tactics by replacing them by a
placeholder, creating a tactic with an unspecified argument.
Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can
try to abstract any of the
\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a variable (playing
the role of a placeholder), we
can replace $u$ by $h$ to create an abstracted tactic.
We denote this intermediate tactic $t[h/u]$. By repeating the
process, multiple
arguments may be abstracted in the same tactic. Ultimately, the more
abstractions are performed the more general a tactic is, as many tactics
become instances of the same abstracted tactic. As a consequence, it becomes
harder and harder to predict suitable arguments. In our experiments, we create
one abstraction $\hat{t}$ for each tactic $t$ by abstracting all subterms of
type theorem list in $t$.

\subsection{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need
to
instantiate the placeholders $h$ inside $\hat{t}$. Because it is difficult to
build a general predictor effective on each type of argument, we manually
design different argument predictors for each type. Those predictors are
given a goal $g$ as input and return the best predicted argument for this
goal. We discuss here specific algorithm for instantiating three common types 
of arguments: theorem list, theorem, term.

\paragraph{Theorem List}
A typical example for

Our default algorithm relies on argument predictions for theorems with a
radius of 16. It uses the produced list
$\mathit{Predictions}^{\mathit{tactic}}_{16}(g)$ to replace all the
placeholders
in $\hat{t}$.
The type of theorem lists is a very common type in \holfour proofs and theorems 
contain enough information to be predicted accurately.
We are helped by the fact that the order of theorems rarely matters in this 
type of tactics and

\paragraph{Theorem}

\paragraph{Term}

We re-use the scheme describe previously for changing the prediction of one 
argument into the prediction of a list of arguments.
The only things that changes is that we have to design a special predictor for 
terms. In this context we think that the predictions should come from the list 
of subterms present in the goal as a first approximation. So all algorithm will 
try

Term synthesis and instantiation is vast problem for higher-order theorems 
provers such as Satallax and inductive programming Hipsec.

Side experiments with independent term argument predictions 
shows that trying to instantiate the tactic with subterms from the goal improve 
the success rate by another 4 percent. 
Nevertheless, in the evaluation presented in this paper, the 
abstraction/instantiation 
method is evaluated only for list of 
theorems, tactics with other type of arguments were predicted with their 
original arguments.

%We have positive results
%with term abstraction but they are not evaluated in
%Section~\ref{sec:evaluation}.

>>>>

\subsection{Abstraction of Tactic Arguments}
The first step is to abstract arguments of tactics by replacing them by a
placeholder, creating a tactic with an unspecified argument.
Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can
try to abstract any of the
\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a variable (playing
the role of a placeholder), we
can replace $u$ by $h$ to create an abstracted tactic.
We denote this intermediate tactic $t[h/u]$. By repeating the
process, multiple
arguments may be abstracted in the same tactic. Ultimately, the more
abstractions are performed the more general a tactic is, as many tactics
become instances of the same abstracted tactic. As a consequence, it becomes
harder and harder to predict suitable arguments.

\subsection{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need
to instantiate the placeholders $h$ inside $\hat{t}$. Because it is difficult to
build a general predictor effective on each type of argument, we manually
design different argument predictors for each type. Those predictors are
given a goal $g$ as input and return the best predicted argument for this
goal. We discuss here specific algorithm for instantiating three common types 
of arguments: theorem list, theorem, term. Creating predictors for other types 
is harder as they lack the structure that is present in mathematical 
formulas.

\paragraph{Theorem List}
The type of theorem lists is the most commonly used type in \holfour proofs.
Here are two examples of such constructions with placeholder 
\texttt{[thm\_list]}: \texttt{REWRITE\_TAC [thm\_list]},  \texttt{METIS\_TAC 
[thm\_list]}.

In general, the list of arguments in such tactics respect the two following 
property:
\begin{itemize}
\item Additional theorems do not cause the tactic to fail and useless theorems 
are ignored. For \texttt{REWRITE\_TAC}, loops are generally avoided because 
most equations appears with only one orientation in the database of theorems.
\item The order of theorems in the list does not influence the final result and 
thus we can consider the list to be a set in our algorithm.
\end{itemize}

Since our predictor is not perfectly accurate and the presented tactics are 
robust, we select a list of theorems usually longer that the original one to 
instantiate these tactics.
Our default predictor for list of theorems selects 16 theorems from the 
database of theorems $\mathbb{D}_{\mathit{theorem}}$. It uses the produced list
$\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= [thm1,\ldots,thm16]$ to 
replace the
placeholders in $\hat{t}$. The instantiation of the first abstracted is thus
\texttt{REWRITE\_TAC [thm1,...,thm16]}.

\paragraph{Theorem}
Instantiating a single theorem is much harder than a list of theorems. Indeed,
the predictions needs to be very precise for the tactic to be effective.
For example, the tactic MPTAC only works if the theorem is an instance
of the precondition. The best option would be to have a train a theorem 
predictor that can adapt its predictions to the given tactic. However, this 
would require a stronger machine learning model than nearest neighbor algorithm.
Therefore, in order to mitigate the gap in the quality of the predictions, we 
propose to create a tactic that tries instantiations until one succeeds. (i.e 
by succeeding we mean produces a list of goal that does not contain the 
original goal)

\paragraph{Term}
We re-use the scheme describe previously for changing the prediction of one 
argument into the prediction of a list of arguments.
The only things that changes is that we have to design a special predictor for 
terms. In this context we think that the predictions should come from the list 
of subterms present in the goal as a first approximation. So all algorithm will 
try

Side experiments with term argument predictions 
show that trying to instantiate the tactic with subterms 
from the goal improve the success rate by another 4 percent. 
However, in the evaluation presented in this paper, the 
abstraction/instantiation method is evaluated only for the type of theorem list.
Moreover, the current argument predictions are only affected by the given goal, 
and we believe it would be beneficial to have more flexible predictors that can 
adapt to their predictions to the abstracted tactic.
